---
title: "OA_monitor_documentation"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Documentation

### Overview

This application provides functionality to retrieve publication records from 
multiple sources (such as ORCID) to provide an overview of the completeness of the respective source
together with the open access status of all publications.

### Data Sources

Currently it is possible to retrieve publications from [ORCID](https://orcid.org/), 
[Pubmed](https://pubmed.ncbi.nlm.nih.gov/), [Google scholar](https://scholar.google.com/) and
[Publons](https://publons.com/search/). 

### Guide 

The following describes a step by step example use case of the app.

#### First step

As a first step, at least one identifier of the author of interest needs to be obtained. 
For each of the following identifiers first go to the respective webpage:

* [ORCID](https://orcid.org/)
* [Google scholar](https://scholar.google.com/)
* [Publons](https://publons.com/search/)

and then search for the author. Navigate to the profile of the researcher 
of interest. Copy the identifier and paste it into the corresponding field in the app 
(alternatively it is possible to directly copy the URL).


For Pubmed a search query has to be specified instead of an identifier. When pressing
'Example Pubmed Query' in the app, a simple example query is generated which needs to 
be adapted (!). For more information on how to write Pubmed queries go to [this help page](https://www.ncbi.nlm.nih.gov/books/NBK3827/#pubmedhelp.How_do_I_search_by_author).

Furthermore there are two additional options that can be selected. The first one is
the possibility to use Crossref for matching publications from Google scholar to other
sources. The use of Crossref is not strictly needed but might find additional matches of publications
from different sources than the standard matching. The drawback of using Crossref is the much longer time it
takes to complete (usually several minutes). It is therefore recommended to first try without
this option and only enable it when the quality of the publication matching is poor (or there are known missing links).
The second additional option is to retrieve citation data from the [NIH Open Citation Collection](https://icite.od.nih.gov/).

#### Second step

After entering at least one identifier, `show report` can be pressed to start retrieval
and merging of the publication data. This will happen in the background and the
app will update as soon as new data becomes available. Depending on the number
of identifiers and the number of publications this can take several minutes.

#### Third step

The obtained data can now be explored using the Histogram on the right hand side and the Table underneath it.

The data can be filtered by using the given selection buttons in the box on the left hand side with the title `Filter options`. 
Filtering options include by dataset, by year and by open access (OA) status. Most useful
for the comparison of the publications in different sources is the option to filter by dataset. 
This follows a logical selection of the datasets, e.g. `In` `Orcid` `and` `In` `Publons` 
(i.e. `orcid & publons` in code) means to only show the publications that are in ORCID and 
in Publons at the same time,  `In` `Orcid` `or` `In` `Publons` would mean all publications
that are in either Publons or ORCID (or both), `Not in` `Orcid` `and` `In` `Publons` 
are the ones that are not in ORCID but at the same time are in Publons.

The Upset plot underneath the filter box shows a 
convenient overview of the different sets that can be found in the data. This plot
is also the main visualization if one is only interested in updating the different sources
since it is immediately visible which sources are complete and which are not.

#### Fourth step

Above the table listing the publications, the button `Bibtex citation` can be pressed to generate a `.tex` citation file
of all the entries in the table. The resulting content will appear underneath the table in a new box and can be downloaded by pressing
`Download Bibtex citation`. This file in turn can be uploaded within the web page of the respective
profile to update the publications in that source.



### Methods

#### Open access status

The open access status is retrieved from [Unpaywall](https://unpaywall.org/) and 
relies on the DOI of a publication. Since Google scholar does not use DOI's it is possible that
the open access status for certain publications is unknown.
Preprints are defined as having `oa == "green"` and `best_oa_location.version == "submittedVersion"` or 
`oa == "green"` and `best_oa_location.version == "publishedVersion"` and `first_oa_location.version == "submittedVersion"` 
(i.e. peer reviewed publication linked in unpaywall).


## About
### Source code

The source code can be found on [www.github.com/markrobinsonuzh/os_monitor](https://github.com/markrobinsonuzh/os_monitor).

### Bug reports

If you encounter problems please open an issue on Github at [www.github.com/markrobinsonuzh/os_monitor/issues](https://github.com/markrobinsonuzh/os_monitor/issues).
Please make a short description of the error you encounter and provide the input that you used (like the ORCID or Google scholar ID). General feedback and feature requests are of course also welcome.

### License

This software is licensed under the [MIT License](https://github.com/markrobinsonuzh/os_monitor/blob/master/LICENSE).

### Contact

If you have other questions do not hesitate to contact us:

Reto Gerber: [reto.gerber@uzh.ch](mailto:reto.gerber@uzh.ch)

Mark D Robinson: [mark.robinson@mls.uzh.ch](mailto:mark.robinson@mls.uzh.ch)

### Session Info

```{r session info}
sessionInfo()
```

