---
title: "Florian Altermatt"
output:
  html_document:
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages/data and set parameters

```{r packages_params_load, message=FALSE}

#rmarkdown::render("analysis/florian_altermatt.Rmd")

library(dplyr)
library(rcrossref)
library(roadoi)
source(here::here("code/functions.R"))

outdir <- here::here("output")
tbl_eprints <- readRDS(file.path(outdir, "tbl_eprints.rds"))
tbl_authorkeys <- readRDS(file.path(outdir, "tbl_authorkeys.rds"))

author <- c("altermatt f 0000 0002 4831 6958", "altermatt f")
orcid <- "0000-0002-4831-6958"
```


# Extract ZORA, ORCID and Unpaywall entries

```{r tables, message=FALSE}
tbl_author <- tbl_authorkeys %>% filter(authorkey %in% author) %>%
  left_join(tbl_eprints)

ws <- retrieve_from_orcid(orcid)

m <- full_join(ws, tbl_author, 
               by="doi", suffix=c(".orcid",".zora"))

oaf <- oadoi_fetch(unique(na.omit(m$doi)),
                   email="mark.robinson@mls.uzh.ch")

m <- m %>% left_join(oaf %>% select(doi, oa_status), 
                     by = "doi", suffix=c(".zora", ".unpaywall"))
```

# Cross-tabulation of ZORA and Unpaywall entries

```{r crosstab, message=FALSE}
with(m, table(oa_status.zora, oa_status.unpaywall, useNA="ifany"))
```


# Articles listed as 'closed' in either ZORA or Unpaywall

```{r}
z <- m %>% filter(oa_status.unpaywall == "closed" | oa_status.zora == "closed") %>% 
  select(doi, eprintid, oa_status.zora, 
         oa_status.unpaywall, title, journal)
z <- z %>% 
  mutate(oa_status.unpaywall = ifelse(is.na(oa_status.unpaywall), "",
                                      paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>"))) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = ifelse(is.na(eprintid), "", paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                                                  eprintid, "'>", eprintid, "</a>")))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```


# Where ZORA and Unpaywall differ in OA Status (assuming they match on DOI)

```{r}
z <- m %>% filter(!is.na(oa_status.unpaywall), 
                  !is.na(oa_status.zora), 
                  oa_status.zora != oa_status.unpaywall) %>% 
  select(doi, eprintid, oa_status.zora, 
         oa_status.unpaywall, title, journal)
z <- z %>% 
  mutate(oa_status.unpaywall = paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>")) %>%
  mutate(doi = paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>")) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>"))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```

# Records in ZORA but not in ORCID (using DOI as key)

```{r}
z <- m %>% filter(!is.na(eprintid), is.na(type.orcid))
z <- z %>% 
  select(doi, eprintid, date, type.zora, refereed, oa_status.zora) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>"))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```

# Records in ORCID but not in ZORA (using DOI as key; type.orcid='other' is usually a preprint)

```{r}
z <- m %>% filter(is.na(eprintid), !is.na(type.orcid))
z <- z %>% 
  select(doi, type.orcid, title, journal) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>")))
DT::datatable(unique(z), extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```


# sessionInfo


```{r pressure, echo=FALSE}
sessionInfo()
```