---
title: "Mark Robinson"
output:
  html_document:
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages/data and set parameters

```{r packages_params_load, message=FALSE}

#rmarkdown::render("analysis/mark_robinson.Rmd")

library(dplyr)
library(rcrossref)
library(roadoi)
library(ggplot2)
source(here::here("code/functions.R"))

outdir <- here::here("output")
tbl_eprints <- readRDS(file.path(outdir, "tbl_eprints.rds"))
tbl_authorkeys <- readRDS(file.path(outdir, "tbl_authorkeys.rds"))

author <- c("robinson m 0000 0002 3048 5518", "robinson m d", "robinson m")
orcid <- "0000-0002-3048-5518"

open_cols <- c("closed" = "gray48", "hybrid" = "darkorange1",
               "green" = "chartreuse4", "gold" = "gold",
               "preprint" = "red", "bronze" = "darkgoldenrod4",
               "blue" = "blue")
my_scale <- scale_fill_manual(values=open_cols)
```


# Extract ZORA, ORCID and Unpaywall entries

```{r tables, message=FALSE}
tbl_author <- tbl_authorkeys %>% 
  filter(authorkey %in% author) %>%
  left_join(tbl_eprints) %>%
  mutate(year = date) %>%
  filter(eprintid != 149052)
dim(tbl_author) # ZORA

# add blue OA
tbl_author$oa_status[tbl_author$published_doc & tbl_author$oa_status=="closed"] <- "blue" 

ws <- retrieve_from_orcid(orcid)
dim(ws) # ORCID

# hacks
# ws$doi[ws$doi=="10.3138/jsp.44-3-005"] <- "10.3138/jsp.44.3.005"
# ws$doi[ws$doi=="10.5061/dryad.8686"] <- "10.1666/10031.1"
# 
# tbl_author$doi[tbl_author$doi=="10.2527/jas2013-6979"] <- "10.2527/jas.2013-6979"
# tbl_author <- tbl_author %>% filter(!(doi %in% c("10.5281/zenodo.60164",
#                                                  "10.19227/jzar.v4i4.194",
#                                                  "10.19227/jzar.v4i4.207",
#                                                  "10.19227/jzar.v5i1.252",
#                                                  "10.19227/jzar.v6i3.355",
#                                                  "10.19227/jzar.v7i2.375")))


m <- full_join(ws, tbl_author, 
               by="doi", suffix=c(".orcid",".zora")) %>%
  filter(doi != "logical(0)")
m$doi <- tolower(m$doi)

oaf <- oadoi_fetch(unique(na.omit(m$doi)),
                   email="mark.robinson@mls.uzh.ch")

m <- m %>% left_join(oaf %>% select(doi, oa_status), 
                     by = "doi", suffix=c(".zora", ".unpaywall")) %>%
  mutate(year = year.orcid)
m$overall_oa <- m$oa_status.unpaywall
m$overall_oa[m$type.orcid=="other"] <- "preprint"
w <- is.na(m$overall_oa)
m$overall_oa[w] <- m$oa_status.zora[w]
w <- is.na(m$year)
m$year[w] <- m$year.zora[w]
w <- m$overall_oa == "closed" & m$oa_status.zora=="blue"
m$overall_oa[w] <- "blue"
```

# Cross-tabulation of ZORA and Unpaywall entries

```{r crosstab, message=FALSE}
with(m, table(oa_status.zora, oa_status.unpaywall, useNA="ifany"))
```


# Plot of OA over time: only ZORA records

```{r plot_oa_zora, message=FALSE}
tbl_author$oa_status <- factor(tbl_author$oa_status, levels = names(open_cols))

ggplot(tbl_author %>% filter(date > 2007, date <= 2020), aes(x=date, fill=oa_status)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90)) +
  my_scale

```

# Plot of OA over time: ZORA + ORCIDs

```{r plot_oa_combined, message=FALSE}
m$overall_oa <- factor(m$overall_oa, levels = names(open_cols))

ggplot(m %>% filter(year > 2005, year <= 2020), aes(x=year, fill=overall_oa)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90)) +
  my_scale
```


# Articles listed as 'closed' in either ZORA or Unpaywall

```{r}
z <- m %>% filter(oa_status.unpaywall == "closed" | oa_status.zora == "closed") %>% 
  select(doi, eprintid, oa_status.zora, 
         oa_status.unpaywall, year.orcid, title, journal) %>%
  arrange(desc(year.orcid))
z <- z %>% 
  mutate(oa_status.unpaywall = ifelse(is.na(oa_status.unpaywall), "",
                                      paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>"))) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = ifelse(is.na(eprintid), "", paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                                                  eprintid, "'>", eprintid, "</a>")))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```


# Where ZORA and Unpaywall differ in OA Status (assuming they match on DOI)

```{r}
z <- m %>% filter(!is.na(oa_status.unpaywall), 
                  !is.na(oa_status.zora), 
                  oa_status.zora != oa_status.unpaywall) %>% 
  select(doi, eprintid, oa_status.zora, year.orcid,
         oa_status.unpaywall, title, journal) %>%
  arrange(desc(year.orcid))

z <- z %>% 
  mutate(oa_status.unpaywall = paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>")) %>%
  mutate(doi = paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>")) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>"))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```

# Records in ZORA but not in ORCID (using DOI as key)

```{r}
z <- m %>% filter(!is.na(eprintid), is.na(type.orcid))
z <- z %>% 
  select(doi, eprintid, type.zora, refereed, oa_status.zora, year.zora) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>")) %>%
  arrange(desc(year.zora))
DT::datatable(z, extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100,
                             scrollY = TRUE, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```

# Records in ORCID but not in ZORA (using DOI as key; type.orcid='other' is usually a preprint)

```{r}
z <- m %>% filter(is.na(eprintid), !is.na(type.orcid))
z <- z %>% 
  select(doi, type.orcid, title, journal, year.orcid) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  arrange(desc(year.orcid))
DT::datatable(unique(z), extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```


# sessionInfo


```{r pressure, echo=FALSE}
sessionInfo()
```