---
title: "Mark Robinson"
output:
  html_document:
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
params:
  pri_author: NA
  sec_author: NA
  orcid: NA
  cutoff_year: NA
  tbl_subjects: NA
  tbl_authorkeys: NA
  tbl_eprints: NA
  unpaywall: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages/data and set parameters

```{r packages_params_load, message=FALSE}
# outdir <- here::here("output")
# tbl_eprints <- readRDS(file.path(outdir, "tbl_eprints.rds"))
# tbl_authorkeys <- readRDS(file.path(outdir, "tbl_authorkeys.rds"))
# tbl_subjects <- readRDS(file.path(outdir, "tbl_subjects.rds"))

# pri_author <- c("robinson m 0000 0002 3048 5518")
# sec_author <- c("robinson m d")
# orcid <- "0000-0002-3048-5518"
# cutoff_year <- 2011

pri_author <- params$pri_author
sec_author <- params$sec_author
orcid <- params$orcid
cutoff_year <- params$cutoff_year
tbl_authorkeys <-  params$tbl_authorkeys
tbl_eprints <-  params$tbl_eprints
tbl_subjects <- params$tbl_subjects
unpaywall <-  params$unpaywall

open_cols <- c("closed" = "gray48", "hybrid" = "darkorange1",
               "green" = "chartreuse4", "gold" = "gold",
               "preprint" = "red", "bronze" = "darkgoldenrod4",
               "blue" = "blue")
my_scale <- scale_fill_manual(values=open_cols)
```


# Extract ZORA, ORCID and Unpaywall entries

```{r tables, message=FALSE}
# tbl_author <- tbl_authorkeys %>% 
#   filter(authorkey %in% c(pri_author, sec_author)) %>%
#   left_join(tbl_eprints) %>%
#   mutate(year = date, doi = tolower(doi))
# tbl_author <- create_tbl_author(tbl_authorkeys,tbl_eprints,pri_author,sec_author)

# dim(tbl_author) # ZORA

# dept_fac <- tbl_author %>% left_join(tbl_subjects %>%
#                                        select(eprintid, name, parent_name))
# institute
# (org_unit <- dept_fac %>% select(name) %>% group_by(name) %>%
#   tally %>% top_n(1) %>% pull(name))
# faculty
# (fac <- dept_fac %>% select(parent_name) %>% group_by(parent_name) %>% 
#   tally %>% top_n(1) %>% pull(parent_name))

zora <- create_zora(pri_author,sec_author,tbl_author,tbl_subjects)
# zora <- dept_fac %>%
#   mutate(dept = name, faculty = parent_name) %>%
#   filter(authorkey == pri_author | (authorkey %in% sec_author & dept == org_unit)) %>%
#   select(-dept, -faculty, -name, -parent_name) %>% unique()
# 
# # add blue OA
# zora$oa_status[zora$published_doc & zora$oa_status=="closed"] <- "blue" 

# ws <- retrieve_from_orcid(orcid) %>%
#   mutate(doi = tolower(doi))
# dim(ws) # ORCID

# hacks
# ws$doi[ws$doi=="10.3390/rs90201010"] <- "10.3390/rs9020110"
# ws$doi <- gsub("doi: ","",ws$doi)
# 
# tbl_author <- tbl_author %>% filter(!(doi %in% c("10.5589/m08-026","10.5589/m08-024")))
# tbl_author$doi[tbl_author$doi=="10.5194/bgd-5-1293-2008, 2008"] <- "10.5194/bgd-5-1293-2008"


# m <- full_join(ws, zora, 
#                by="doi", suffix=c(".orcid",".zora")) %>%
#   filter(doi != "logical(0)")

# oaf <- oadoi_fetch(unique(na.omit(m$doi)),
#                    email="mark.robinson@mls.uzh.ch")
# oaf <- oadoi_fetch_local(unique(na.omit(m$doi)),unpaywall)

# m <- create_combined_data(ws,zora,oaf)
# m <- m %>% left_join(oaf %>% select(doi, oa_status), 
#                      by = "doi", suffix=c(".zora", ".unpaywall")) %>%
#   mutate(year = year.orcid, title = title.orcid)
# m$overall_oa <- m$oa_status.unpaywall
# m$overall_oa[m$type.orcid=="other"] <- "preprint"
# w <- is.na(m$overall_oa)
# m$overall_oa[w] <- m$oa_status.zora[w]
# w <- is.na(m$year)
# m$year[w] <- m$year.zora[w]
# w <- is.na(m$title)
# m$title[w] <- m$title.zora[w]
# w <- m$overall_oa == "closed" & m$oa_status.zora=="blue"
# m$overall_oa[w] <- "blue"

m <- create_combined_data_wrapper(tbl_authorkeys,tbl_eprints,tbl_subjects,pri_author,sec_author,orcid,unpaywall)



```

# Cross-tabulation of ZORA and Unpaywall entries

```{r crosstab, message=FALSE}
with(m, table(oa_status.zora, oa_status.unpaywall, useNA="ifany"))
```


# Plot of OA over time: only ZORA records

```{r plot_oa_zora, message=FALSE, fig.width=14}

ggplot(zora %>% filter(date >= cutoff_year, date <= 2020), aes(x=date, fill=oa_status)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("ZORA OA Status") +
  my_scale
```

# Articles listed as 'closed' in ZORA

```{r}
z <- zora %>% filter(oa_status == "closed") %>% 
  select(doi, eprintid, type, refereed, title, oa_status, year) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>")) %>%
  arrange(desc(year))
DT::datatable(z, extensions = 'Buttons',
              options = list(dom = 'Bfrtip',
                             pageLength = 200,
                             buttons = list('copy', 'csv', 'excel')), 
              escape = FALSE, rownames = FALSE)
```


# Plot of OA over time: ZORA + ORCIDs

```{r plot_oa_combined, message=FALSE, fig.width=14}
# m$overall_oa <- factor(m$overall_oa, levels = names(open_cols))

ggplot(m %>% filter(year >= cutoff_year, year <= 2020), aes(x=year, fill=overall_oa)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("ZORA + ORCID OA Status") +
  my_scale
```


# OA percent over time (w/ or w/o blue as OA)

```{r blue}
z <- m %>% filter(year >= cutoff_year) %>% 
  group_by(year) %>% 
  summarize(oa_pct_blue = round(100*mean(overall_oa != "closed"),2),
            oa_pct = round(100*mean(!(overall_oa %in% c("blue","closed")))),
            closed = sum(overall_oa == "closed")) %>% 
  data.frame()
DT::datatable(z, extensions = 'Buttons',
              options = list(dom = 'Bfrtip',
                             pageLength = 200,
                             buttons = list('copy', 'csv', 'excel')), 
              escape = FALSE, rownames = FALSE)
```

# Articles listed as 'closed' overall

```{r}
z <- m %>% filter(overall_oa == "closed", year >= cutoff_year) %>% 
  select(doi, eprintid, overall_oa, oa_status.zora, 
         oa_status.unpaywall, year, title, journal) %>%
  arrange(desc(year))
z <- z %>% 
  mutate(oa_status.unpaywall = ifelse(is.na(oa_status.unpaywall), "",
                                      paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>"))) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = ifelse(is.na(eprintid), "", paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                                                  eprintid, "'>", eprintid, "</a>")))
DT::datatable(z, extensions = 'Buttons',
              options = list(dom = 'Bfrtip',
                             pageLength = 200,
                             buttons = list('copy', 'csv', 'excel')), 
              escape = FALSE, rownames = FALSE)
```


# Records in ZORA but not in ORCID (using DOI as key)

```{r}
z <- m %>% filter(!is.na(eprintid), is.na(type.orcid), year >= cutoff_year)
z <- z %>% 
  select(doi, eprintid, type.zora, refereed, title, oa_status.zora, year) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>")) %>%
  arrange(desc(year))
DT::datatable(z, extensions = 'Buttons',
              options = list(dom = 'Bfrtip',
                             pageLength = 200,
                             buttons = list('copy', 'csv', 'excel')), 
              escape = FALSE, rownames = FALSE)
```

# Records in ORCID but not in ZORA (using DOI as key; excludes type.orcid='other', which is usually a preprint)

```{r}
z <- m %>% filter(is.na(eprintid), !is.na(type.orcid), 
                  year >= cutoff_year, type.orcid != "other")
z <- z %>% 
  select(doi, type.orcid, title, journal, year) %>%
  mutate(doi = ifelse(is.na(doi), "", paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>"))) %>%
  arrange(desc(year))
DT::datatable(unique(z), extensions = list('Buttons','Scroller','FixedHeader'),
              class = 'cell-border stripe',
              options = list(dom = 'Bfrtip',
                             buttons = c('excel', "csv"),
                             pageLength = 100, fixedHeader = TRUE),
              escape = FALSE, rownames = FALSE)
```

# Where ZORA and Unpaywall differ in OA Status (assuming they match on DOI)

```{r}
z <- m %>% filter(!is.na(oa_status.unpaywall), 
                  !is.na(oa_status.zora), 
                  oa_status.zora != oa_status.unpaywall,
                  year >= cutoff_year) %>% 
  select(doi, eprintid, oa_status.zora, year,
         oa_status.unpaywall, title, journal) %>%
  arrange(desc(year))

z <- z %>% 
  mutate(oa_status.unpaywall = paste0("<a href='https://api.unpaywall.org/v2/",
                                      doi,"?email=YOUR_EMAIL'>",
                                      oa_status.unpaywall, "</a>")) %>%
  mutate(doi = paste0("<a href='https://www.doi.org/",
                      doi, "'>", doi, "</a>")) %>%
  mutate(eprintid = paste0("<a href='https://www.zora.uzh.ch/id/eprint/",
                           eprintid, "'>", eprintid, "</a>"))
DT::datatable(z, extensions = 'Buttons',
              options = list(dom = 'Bfrtip',
                             pageLength = 200,
                             buttons = list('copy', 'csv', 'excel')), 
              escape = FALSE, rownames = FALSE)
```

# sessionInfo

```{r pressure, echo=FALSE}
sessionInfo()
```