---
title: "Report"
output:
  html_document:
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
params:
  pri_author: NA
  sec_author: NA
  orcid: NA
  cutoff_year: NA
  tbl_subjects: NA
  tbl_authorkeys: NA
  tbl_eprints: NA
  unpaywall: NA
  zora: NA
  m: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages/data and set parameters

```{r packages_params_load, message=FALSE}
# outdir <- here::here("output")
# tbl_eprints <- readRDS(file.path(outdir, "tbl_eprints.rds"))
# tbl_authorkeys <- readRDS(file.path(outdir, "tbl_authorkeys.rds"))
# tbl_subjects <- readRDS(file.path(outdir, "tbl_subjects.rds"))

# pri_author <- c("robinson m 0000 0002 3048 5518")
# sec_author <- c("robinson m d")
# orcid <- "0000-0002-3048-5518"
# cutoff_year <- 2011

pri_author <- params$pri_author
sec_author <- params$sec_author
orcid <- params$orcid
cutoff_year <- params$cutoff_year
tbl_authorkeys <-  params$tbl_authorkeys
tbl_eprints <-  params$tbl_eprints
tbl_subjects <- params$tbl_subjects
unpaywall <-  params$unpaywall
zora <- params$zora
m <- params$m


open_cols <- c("closed" = "gray48", "hybrid" = "darkorange1",
               "green" = "chartreuse4", "gold" = "gold",
               "preprint" = "red", "bronze" = "darkgoldenrod4",
               "blue" = "blue")
my_scale <- scale_fill_manual(values=open_cols)
```


# Cross-tabulation of ZORA and Unpaywall entries

```{r crosstab, message=FALSE}
with(m, table(oa_status.zora, oa_status.unpaywall, useNA="ifany"))
```


# Plot of OA over time: only ZORA records

```{r plot_oa_zora, message=FALSE, fig.width=14}
zora_only_plot(zora,cutoff_year)
```

# Articles listed as 'closed' in ZORA

```{r}
closed_in_zora_table(zora)
```


# Plot of OA over time: ZORA + ORCIDs

```{r plot_oa_combined, message=FALSE, fig.width=14}
zora_orcid_plot(m,cutoff_year)
```


# OA percent over time (w/ or w/o blue as OA)

```{r blue}
oa_percent_time_table(m,cutoff_year)
```

# Articles listed as 'closed' overall

```{r}
overall_closed_table(m,cutoff_year)
```


# Records in ZORA but not in ORCID (using DOI as key)

```{r}
zora_without_orcid_table(m,cutoff_year)
```

# Records in ORCID but not in ZORA (using DOI as key; excludes type.orcid='other', which is usually a preprint)

```{r}
orcid_without_zora_table(m,cutoff_year)
```

# Where ZORA and Unpaywall differ in OA Status (assuming they match on DOI)

```{r}
oa_status_diff_zora_unpaywall_table(m,cutoff_year)
```

# sessionInfo

```{r pressure, echo=FALSE}
sessionInfo()
```